### 2021-08-25 博士/论文提交/adversarial_attack/TCSVT/latex/coverletter.tex
删除正文中的下划线，写cover letter。

\documentstyle{letter}
\addtolength{\voffset}{-0.5in}
\addtolength{\hoffset}{-0.3in}
\addtolength{\textheight}{2cm}
\begin{document}
\signature{Zhenbang Li, Yaya Shi, Jin Gao$^*$, Shaoru Wang, Bing Li, Pengpeng Liang, Weiming Hu}

\begin{letter}{}
\opening{\textbf{Dear Editors,}}

We would like to express our heartfelt gratitude to you and the reviewers for the insightful and helpful comments.
The editor-in-chief suggested us to revise the manuscript ``TCSVT-06387-2021: A Simple and Strong Baseline for Universal Targeted Attacks on Siamese Visual Tracking" and resubmit it to TCSVT.
When we revised the paper, we carefully considered and followed all the comments and suggestions provided by you and the reviewers. To summarize, we have made the following revisions:

(1) We have replaced the phase ``imperceptible perturbation" with ``translucent perturbation" in the manuscript to avoid the ambiguity.

(2) We have carefully considered and followed all the comments and suggestions related to the clarity of the writing, and made the new material a thoroughly revised manuscript.

(3) We have summarized the characteristics of the used datasets in our experiments.

(4) We have compared with state-of-the-art attack methods including RTAA, SPARK, TTP, CSA and FAN in both targeted and untargeted attack settings.

(5) We have added experiments to train perturbations in the absence of ground truth box information.

We hope that our revised manuscript is now appropriate for publication in IEEE Transactions on Circuits and Systems for Video Technology. Specific responses to all the comments of each reviewer are included in the rest of this document and highlighted using bold font after the comments of each reviewer for the convenience of cross-reference. To make the changes easier to identify where necessary, we also have underlined most of the revised parts in the manuscript and provide an underlined version for the convenience of second review.

\vspace{2\parskip}
\closing{Yours Sincerely,}
\vspace{5mm}
$^*$ Corresponding author:
\vspace{-2mm}

Name: Jin Gao   \quad\quad  E-mail: jin.gao@nlpr.ia.ac.cn
\end{letter}

\end{document}