### 2021-09-29 博士/毕业论文/ucasthesis/Tex/Backmatter_blind.tex

%---------------------------------------------------------------------------%
%->> Backmatter
%---------------------------------------------------------------------------%
\chapter{作者攻读学位期间发表的学术论文与研究成果}

\section*{已发表（或正式接受）的学术论文:}

{
\setlist[enumerate]{}% restore default behavior
\begin{enumerate}[nosep]
    \item "Globally Spatial-Temporal Perception: a Long-Term Tracking System", \textit{International Conference on Image Processing} (\textbf{ICIP}), 2020. [本人第一作者]
    \item "End-to-End Temporal Feature Aggregation for Siamese Trackers", \textit{International Conference on Image Processing} (\textbf{ICIP}), 2020. [本人第一作者]
    \item "Manipulating Template Pixels for Model Adaptation of Siamese Visual Tracking", \textit{IEEE Signal Processing Letters}, 2020. [本人第一作者]
\end{enumerate}
}

\section*{已投稿的学术论文:}

{
\setlist[enumerate]{}% restore default behavior
\begin{enumerate}[nosep]
    \item "A Simple and Strong Baseline for Universal Targeted Attacks on Siamese Visual Tracking", \textit{IEEE Transactions on Circuits and Systems for Video Technology}, 2021. [本人第一作者]（Major Revision，已修改并提交）
    \item "Learning Video-Text Aligned Representations for Video Captioning", \textit{Thirty-Sixth AAAI Conference on Artificial Intelligence} (\textbf{AAAI}), 2022. [本人第三作者]
\end{enumerate}
}

\chapter[致谢]{致\quad 谢}\chaptermark{致\quad 谢}% syntax: \chapter[目录]{标题}\chaptermark{页眉}
\thispagestyle{noheaderstyle}% 如果需要移除当前页的页眉

\cleardoublepage[plain]% 让文档总是结束于偶数页，可根据需要设定页眉页脚样式，如 [noheaderstyle]
%---------------------------------------------------------------------------%